\change{The effect of the COVID-19 pandemic on arXiv pre-prints in maths, physics, computer science, and quantitative biology}

### 2022-10-28 article/ms.tex
minor

The COVID-19 pandemic appears to have had minimal impact on collective research productivity to date, \change{and that is before accounting for the tremendous work in scientific communication and contributions to public policy throughout the pandemic.} However, some significant changes were noted within specific areas of research. The largest persistent decline in pre-prints has occurred in {general physics (arXiv primary category \texttt{physics}), largely due to a decline in  pre-prints posted to computational and applied physics areas within the \texttt{physics} primary category. As of mid-2022, this steady decline continues.} Amongst individual fields the largest \emph{relative} decline in a single field is explainable by a high energy lattice physics conference that was cancelled in 2020. In this study we have only focussed on the quantity of \change{pre-prints} as a measure for productivity, and not their quality. While citations are the most used measure of impact of academic publications, that metric becomes a more biased statistic when many related pre-prints are all being posted nearly at the same time\cite{Fassin:2021}. The relatively long timescales of academic research would suggest that the full impact on academic productivity due to the COVID-19 pandemic is yet to be seen. The methodology and results presented here provide a quantitative framework for future investigations of the longer term impacts of the pandemic.

@@

\change{There were 234 pre-prints posted to the \texttt{q-bio} primary category in 2020 which include pandemic-related terms, and where no author had never appeared on the \texttt{q-bio} pre-print before. We investigated every pre-print to extract the affiliation of the lead author, and used this information to search the internet for the author's website. From this information we identified their research specialisations using their own words, or from their social media pages if no website could be found. In situations where an author's specialisation crossed multiple fields (e.g., mathematics and biology) we took a conservative approach and described the author as a biologist. However, we found that only 3\% (8/234) were led by established biologists who had not appeared in \texttt{q-bio} pre-prints before. About half (45\%; 107/234) were led by mathematicians or physicists, with another 26\% (61/234) led by engineers, computer scientists, or economists.}

@@

>>>>

\change{The COVID-19 pandemic appears to have had minimal impact on collective research productivity {in the science and technology specialisations we examined}, {and that is before accounting for the tremendous work in scientific communication and contributions to public policy throughout the pandemic.} These findings are at odds with similar studies in fields that have larger research overlap with the SARS-CoV-2 virus. For example, publications in life sciences saw a 6.5 fold increase in subject headings related to COVID-19\cite{Riccaboni}. The effects of the pandemic clearly impact research specialisations differently, and we note some} significant changes within specific research areas. The largest persistent decline in pre-prints has occurred in {general physics (arXiv primary category \texttt{physics}), largely due to a decline in  pre-prints posted to computational and applied physics areas within the \texttt{physics} primary category. As of mid-2022, this steady decline continues.} Amongst individual fields the largest \emph{relative} decline in a single field is explainable by a high energy lattice physics conference that was cancelled in 2020.

In this study we have only focussed on the quantity of \change{pre-prints} as a measure for productivity, and not their quality. While citations are the most used measure of impact of academic publications, that metric becomes a more biased statistic when many related pre-prints are all being posted nearly at the same time\cite{Fassin:2021}. The relatively long timescales of academic research would suggest that the full impact on academic productivity due to the COVID-19 pandemic is yet to be seen. The methodology and results presented here provide a quantitative framework for future investigations of the longer term impacts of the pandemic.

@@

\change{There were 234 pre-prints posted to the \texttt{q-bio} primary category in 2020 which include pandemic-related terms, and where no author had ever appeared on the \texttt{q-bio} pre-print before. We investigated every pre-print to extract the affiliation of the lead author, and used this information to search the internet for the author's website. From this information we identified their research specialisations using their own words, or from their social media pages if no website could be found. In situations where an author's specialisation crossed multiple fields (e.g., mathematics and biology) we took a conservative approach and described the author as a biologist. However, we found that only 3\% (8/234) were led by established biologists who had not appeared in \texttt{q-bio} pre-prints before. About half (45\%; 107/234) were led by mathematicians or physicists, with another 26\% (61/234) led by engineers, computer scientists, or economists.}

@@

\bibitem{Riccaboni}
Riccaboni, M., Verginer, L. {The impact of the COVID-19 pandemic on scientific research in the life sciences}, \emph{PLOS One}, \textbf{17}, 2, 1--16, doi:10.1371/journal.pone.0263001 \newblock (2022).
